# The fatal impact of tornadoes and economic effects of floods #

## Synopsis ##

Get the data from NOAA. Analyse impact by deaths. Tornadoes come first. Analyse impact by economy. Floods come first.

## Data Processing ##

### Load ###

How the data were loaded into R

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(fileUrl, destfile = "tempdata.csv.bz2", method = "curl")
stormdata <- read.csv("./tempdata.csv.bz2")
```

### Process ###

How the data are processed for analysis

```{r}
library(Hmisc)
library(reshape)
library(ggplot2)
library(car)

stormdata$EVTYPE <- capitalize(tolower(stormdata$EVTYPE))

damages<-aggregate(cbind(FATALITIES, INJURIES) ~ EVTYPE , stormdata, sum)
dam<-melt(head(damages[order(-damages$FATALITIES,-damages$INJURIES),],10))

stormdata$PROPDMG<-stormdata$PROPDMG*as.numeric(Recode(stormdata$PROPDMGEXP, "'0'=1;'1'=10;'2'=100;'3'=1000;'4'=10000;'5'=100000;'6'=1000000;'7'=10000000;'8'=100000000;'B'=1000000000;'h'=100;'H'=100;'K'=1000;'m'=1000000;'M'=1000000;'-'=0;'?'=0;'+'=0",as.factor.result=FALSE))
stormdata$CROPDMG<-stormdata$CROPDMG*as.numeric(Recode(stormdata$CROPDMGEXP, "'0'=1;'2'=100;'B'=1000000000;'k'=1000;'K'=1000;'m'=1000000;'M'=1000000;''=0;'?'=0",as.factor.result=FALSE))

economic<-aggregate(cbind(PROPDMG, CROPDMG) ~ EVTYPE , stormdata, sum)
eco<-melt(head(economic[order(-economic$PROPDMG,-economic$CROPDMG),],10))
```

## Results ##

### Human casualties ###

Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

```{r}
ggplot(dam, aes(x=EVTYPE,y=value,fill=variable)) + geom_bar(stat = "identity") + coord_flip() +
  ggtitle("Harmful events") + labs(x = "", y="number of people impacted") +
  scale_fill_manual (values=c("#555555","#888888"), labels=c("Deaths","Injuries"))
```

### Economic impact ###

Across the United States, which types of events have the greatest economic consequences?

```{r}
ggplot(eco, aes(x=EVTYPE,y=value,fill=variable)) + geom_bar(stat = "identity") + coord_flip() +
  ggtitle("Economic consequences") + labs(x = "", y="cost of damages in dollars") +
  scale_fill_manual (values=c("#555555","#888888"), labels=c("Property Damage","Crop Damage"))
```